\SW: I. Crowd-sourcing the Discovery of Gravitational Lenses

### 2015-04-21 doc/sw-system.tex
Addressing Anu's comments, done Sections 5.2 and 5.3 but not 5.1 #112

To investigate the properties of the \sw crowd, as characterised by their
agents, we define the following quantities and plot their one and 2-D marginal
distributions in
\Fref{fig:crowd:cornerplot}. This figure only shows the \StageOne agents for clarity,
but the trends we found the same trends in \StageTwo.
%
\begin{description}
%
\item{\noindent\bf ``Effort:''} The number of test images, $\effort$, classified
by a volunteer. In \StageOne, the mean effort per agent was 263; in the shorter
\StageTwo it was 81.
%
\item{\noindent\bf ``Experience:''} The number of training images, $\experience$,
classified by a volunteer. In \StageOne, the mean experience per agent was 29;
in \StageTwo (where the training image frequency was set higher) it was 34.
%
\item{\noindent\bf ``Skill:''} The expectation value of the information gained
per classification (in bits) by that volunteer,  $\skill$, for subjects which
have lens probability 0.5 (\Aref{appendix:swap}). Random classifiers have
$\skill = 0.0$~bits, while perfect classifiers have $\skill = 1.0$~bit.
All agents start
with $\skill = 0.0$~bits. The skill of  an agent increases as training subjects are
classified, and the agent's estimates of its confusion matrix elements improve.
In \StageOne, the mean skill per agent was 0.04 bits; in  \StageTwo it was
0.05~bits.
%
\item{\noindent\bf ``Contribution:''} The integrated skill over a volunteer's
{\it test subject} classification history, a quantity representative of the
total contribution to the project made by that volunteer (see the appendix for
more discussion of this quantity). The classifications of training images
determine the skill, while the classification of test images determines
contribution.  In \StageOne, the mean  contribution per agent was
34.9 bits; in \StageTwo it was 33.5.
%
\item{\noindent\bf ``Information:''} The actual total information $\information$
generated by the agent during the volunteer's classification activity. This
quantity depends on the value of each subject's lens probability when that
subject was presented to the volunteer (simply because information gain is
defined in terms of posterior relative to prior probabilities,
\Aref{appendix:swap}).  Because of this, we expect this quantity to be noisier
than the ``contribution'' defined above.
%
\end{description}

@@

At high values of experience and effort, the skill is \emph{always high}. There
seem to be very few agents logging large numbers of classifications at low skill
(although there are one or two exceptions): almost all high effort
``super-users'' have high skill. These two properties are reflected in both the
contributions these volunteers make (third row) and the information they
generate (fourth row).

The distributions for the \StageTwo agents (orange) are qualitatively similar to
those for the \StageOne agents (blue). Differences are: 1) the maximum effort
possible at \StageTwo is smaller, because fewer subjects were available to be
classifed, but 2) the mean effort expended at \StageTwo was higher (perhaps
because the subjects were higher probability, and more compelling); 3) the
information generated per agent was higher at \StageTwo, because the subjects
had higher probability.

@@

\Fref{fig:crowd:cornerplot} shows the \SW crowd to have quite broad
distributions of logarithmic effort, skill, and contribution. To better
quantify the contributions made by the volunteers, we show their cumulative
distribution on a linear scale in \Fref{fig:crowd:cumulplot}. This plot shows
clearly the importance of the hardest-working, most active volunteers: at
\StageOne,  1.0\% of the volunteers -- 375 people -- made 90\% of the
contribution.  At \StageTwo, where it was not possible to make as
many classifications before running out of subjects, 7.2\% of the volunteers
-- 141 people -- made 90\% of the contribution.

However, it is not the case that only these small groups were {\it capable} of
making large contributions. The cumulative distribution of agent skill is shown
in \Fref{fig:crowd:cumulskillplot}: these distributions are significantly
broader than the corresponding distributions of agent contribution, in
\Fref{fig:crowd:cumulplot}. The most skilled 20\% of agents only possess around
80\% of the skill at each stage. The inexperienced volunteers also possess a
significant fraction of the skill: the most skillful 20\% of experienced
volunteers (1824 people)  possess just 43\% of the total skill. The breadth of
the skill distributions suggests that  the high level of contribution made at
\SW by experienced volunteers is largely a matter of choice (or perhaps,
availability of time!). There were many other volunteers that  were skillful
enough to make large contributions -- they just didn't classify as many images.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering\includegraphics[width=0.9\linewidth]{sw-system-figs/crowd_skill_cumul.pdf}
\caption{Broad cumulative distributions of agent skill:
the most skilled 20\% of the crowd only possess 79\% of the skill at \StageOne.
The \StageOne agents
are shown in blue, the \StageTwo agents in orange. ``Experienced volunteers''
classified 10 or more training subjects.}
\label{fig:crowd:cumulskillplot}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

How fast does the \SW crowd classify subjects? Each agent records the timestamp
of each classification its volunteer makes; by measuring the time lags between
successive classifications we can make estimates of the crowd's classification
and contribution speed. We plot these two quantities for both \StageOne and
\StageTwo in \Fref{fig:results:speed}, normalizing the speed and time axes to
the respective totals. The fractional classification rates in each stage of the
survey fall off in approximately the same way, despite the factor of nearly 20
difference in crowd size. \StageOne (consisting of $\sim430,000$ subjects) was
completed in around 5000 hours (by some 37,000 participants), with
classification rates in the first few days averaging at $\sim10^4$ per hour.
This was stimulated by various forms of advertising (press releases, and emails
to registered Zooniverse users). The asymptotic classification rate was  around
ten times lower, at $\sim10^3$ per hour. Interestingly, the average skill was
found to be approximately constant over the lifetime of \StageOne, leading to a
contribution rate that tracks closely the classification rate. (A 10\% increase
in average skill was seen over the first 40 days, and a decrease of the same
amount in the last 40 days -- not enough to cause a significant difference
between classification and contribution rate behavior, but perhaps reflecting a
learning period at the start and a decrease in participation late on.) \StageTwo
(3700 subjects) was completed in around 600 hours by around 2000 participants,
with classification rates starting at $~\sim2000$ per hour and decaying to
$\sim100$ per hour.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering\includegraphics[width=\linewidth]{sw-system-figs/rate_classification.pdf}
\caption{Fractional classification rate in the \SW \cfhtls survey.}
\label{fig:results:speed}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The similarities between \StageOne and 2, suggest that these numbers could be
scaled to estimate the speed of future \SW projects (or at least those following
a similar blind search strategy). For example, the completion time for a \SW
project may be approximated as
\begin{equation}
    \tau \approx 18\,{\rm days} \left(\frac{10^4}{N_p}\right)
                               \left(\frac{N_s}{10^4}\right),
\label{eq:speed}
\end{equation}
where $N_s$ is the number of subjects to be classified, and $N_p$ is the number
of volunteers in the crowd.

@@

If the system efficiency could be improved by a factor of three, we would
expect the coefficient in \Eref{eq:speed} to decrease by roughly the same factor,
suggesting that an dataset of $10^4$~images could be searched for lenses by a
crowd of $10^4$ volunteers in about 6 days.

>>>>

To investigate the properties of the \sw crowd, as characterised by their
agents, we define the following quantities and plot their one and 2-D marginal
distributions in
\Fref{fig:crowd:cornerplot}. This figure only shows the \StageOne agents for clarity,
but the trends we found the same trends in \StageTwo.
%
\begin{description}
%
\item{\noindent\bf ``Effort:''} The number of test images, $\effort$, classified
by a volunteer. In \StageOne, the mean effort per agent was 263; in the shorter
\StageTwo it was 81.
%
\item{\noindent\bf ``Experience:''} The number of training images, $\experience$,
classified by a volunteer. In \StageOne, the mean experience per agent was 29;
in \StageTwo (where the training image frequency was set higher) it was 34.
%
\item{\noindent\bf ``Skill:''} The expectation value of the information gained
per classification (in bits) by that volunteer,  $\skill$, for subjects which
have lens probability 0.5 (\Aref{appendix:swap}). Random classifiers have
$\skill = 0.0$~bits, while perfect classifiers have $\skill = 1.0$~bit.
All agents start
with $\skill = 0.0$~bits. The skill of  an agent increases as training subjects are
classified, and the agent's estimates of its confusion matrix elements improve.
In \StageOne, the mean skill per agent was 0.04 bits; in  \StageTwo it was
0.05~bits.
%
\item{\noindent\bf ``Contribution:''} This is the integrated skill over a volunteer's
{\it test subject} classification history, a quantity representative of the
total contribution to the project made by that volunteer (see the appendix for
more discussion of this quantity). The classifications of training images
determine the skill, while the classification of test images determines
contribution.  In \StageOne, the mean  contribution per agent was
34.9 bits; in \StageTwo it was 33.5.
%
\item{\noindent\bf ``Information:''} The total information $\information$
generated by the agent during the volunteer's classification activity. This
quantity depends on the value of each subject's lens probability when that
subject was presented to the volunteer (simply because information gain is
defined in terms of posterior relative to prior probabilities,
\Aref{appendix:swap}).  Because of this, we expect this raw quantity to be noisier
than the ``contribution'' defined above. The lower righthand panel of
\Fref{fig:crowd:cornerplot} confirms this: there is a strong, linear correlation
between contribution and information gain, but at a given contribution, the
distribution of information gain has a tail at high values, corresponding to
volunteers that were lucky in the number of high probability subjects that they
happened to be shown (the information gain per classification is maximized when the
subject probability is 0.5, while most subjects have probabilities closer to or less than the prior).
In most of the following discussion we use the contribution when characterizing
crowd participation, but we include
the information gain for completeness.
%
\end{description}

@@

At high values of experience and effort, the skill is \emph{always high}. There
seem to be very few agents logging large numbers of classifications at low
skill: {\it almost all high effort
``super-users'' have high skill.} These two properties are reflected in both the
contributions these volunteers make (third row) and the information they
generate (fourth row).

We found that the distributions for the \StageTwo agents to be qualitatively
very similar to those for the \StageOne agents. The differences are: 1) the
maximum effort possible at \StageTwo is smaller, simply because fewer subjects
were available to be classifed, and 2) the information generated per agent was
slightly higher at \StageTwo, just because the subjects had (on average) higher
probability.

@@

\Fref{fig:crowd:cornerplot} shows the \SW crowd to have quite broad
distributions of logarithmic effort, skill, and contribution. To better
quantify the contributions made by the volunteers, we show their cumulative
distribution on a linear scale in \Fref{fig:crowd:cumulplot}. This plot shows
clearly the importance of the most active volunteers: at
\StageOne,  1.0\% of the volunteers -- 375 people -- made 90\% of the
contribution.  At \StageTwo, where it was not possible to make as
many classifications before running out of subjects, 7.2\% of the volunteers
-- 141 people -- made 90\% of the contribution.

However, it is not the case that only these small groups were {\it capable} of
making large contributions. The cumulative distribution of agent skill is shown
in \Fref{fig:crowd:cumulskillplot}: these distributions are significantly
broader than the corresponding distributions of agent contribution, in
\Fref{fig:crowd:cumulplot}. For example,
the most skilled 20\% of agents possess around 80\% of the skill.

The inexperienced volunteers also possess a significant fraction of the skill.
We find that dividing the crowd into ``experienced volunteers'' (who have seen
10 or more training images), and ``inexperienced volunteers'' (the rest),
results in two groups containing approximately equal total skill (see the dashed
curve in \Fref{fig:crowd:cumulskillplot}). The most skillful 20\% of {\it
experienced} volunteers (1824 people) only possess 43\% of the total skill.
The breadth of the skill distributions suggests that  the high level of
contribution made at \SW by experienced volunteers is largely a matter of choice
(or perhaps, availability of time!). There were many other volunteers that  were
skillful enough to make large contributions -- they just didn't classify as many
images.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering\includegraphics[width=0.9\linewidth]{sw-system-figs/crowd_skill_cumul.pdf}
\caption{Broad cumulative distributions of agent skill:
the most skilled 20\% of the crowd only possess around 80\% of the skill,
at \StageOne. The \StageOne agents
are shown in blue, the \StageTwo agents in orange. ``Experienced volunteers''
classified 10 or more training subjects.}
\label{fig:crowd:cumulskillplot}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

How fast does the \SW crowd classify subjects? Each agent records the timestamp
of each classification its volunteer makes; by measuring the time lags between
successive classifications, we can make estimates of the crowd's classification
and contribution speed. We plot the former quantity for both \StageOne and
\StageTwo in \Fref{fig:results:speed}, normalizing the speed and time axes to
their respective totals. The fractional classification rates
in each stage of the survey fall off in approximately the same way, despite the
factor of nearly 20 difference in crowd size. \StageOne (consisting of
$\sim430,000$ subjects) was completed in around 5000 hours (by some 37,000
participants), with classification rates in the first few days averaging at
$\sim10^4$ per hour. This was stimulated by various forms of advertising (press
releases, and emails to registered Zooniverse users). The asymptotic
classification rate was  around ten times lower, at $\sim10^3$ per hour.
Interestingly, the average skill was found to be approximately constant over the
lifetime of \StageOne, leading to a contribution rate that tracks closely the
classification rate. (A 10\% increase in average skill was seen over the first
40 days, and a decrease of the same amount in the last 40 days -- not enough to
cause a significant difference between classification and contribution rate
behavior, but perhaps reflecting a learning period at the start and then a
decrease
in participation late on.) \StageTwo (3700 subjects) was completed in around 600
hours by around 2000 participants, with classification rates starting at
$~\sim2000$ per hour and decaying to $\sim100$ per hour.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering\includegraphics[width=\linewidth]{sw-system-figs/rate_classification.pdf}
\caption{Fractional classification rate in the \SW \cfhtls survey. Fractional classification rate is the number
of classifications per unit time, divided by the final total number of
classifications. Fractional survey time is the time elapsed since the beginning
of the survey, divided by the total length of the survey. The unit of time in
the fractional  classification rate is one survey, so that the curves in
integrals under the curves are both one. }
\label{fig:results:speed}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The similarities between \StageOne and 2, suggest that these numbers could be
scaled to estimate the speed of future \SW projects (or at least those following
a similar blind search strategy). For example, the completion time for a \SW
project may be approximated as
\begin{equation}
    \tau \approx \tau_0 \left(\frac{10^4}{N_p}\right)
                               \left(\frac{N_s}{10^4}\right),
\label{eq:speed}
\end{equation}
where $N_s$ is the number of subjects to be classified, and $N_p$ is the number
of volunteers in the crowd. Just using the numbers above, we find a characteristic
timescale of $\tau_0 = 18\,{\rm days}$, but as we will see in the next section,
we expect this to be significantly shorter in future projects.

@@

If the system efficiency could be improved by a factor of three, we would expect
the characteristic completion time $\tau_0$  in \Eref{eq:speed} to decrease by
roughly the same factor, suggesting that a dataset of $10^4$~images could be
searched for lenses by a crowd of $10^4$ volunteers in about 6 days.
